---
title: "Getting Figures and Tables Presentation Ready"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

This markdown will go through a few examples of presentation ready tables and figures versus non presentation ready tables and figures. Often, to get tables and figures to a presentation ready form, it only requires a bit more code or even just one more function call. 

```{r, echo = FALSE}
# Load in libraries
library(tidyverse)
library(stringr)
library(knitr)
library(broom)

# Get specific date's historical TMINs and TMAXs
GetDayTemp = function(data) {
  current_day = format(Sys.Date(), "%m-%d")
  filtered_data = data %>%
    filter(grepl(current_day, date))
  return(filtered_data)
}


# Read in temperature data for St. Paul and Urbana
stpaul_data = read.csv("Data/stpaul-temps.csv")

urbana_data = read.csv("Data/urbana-temps.csv")

# Combine St. Paul and Urbana data in a list to use lapply functions
data_list = list(stpaul_data, urbana_data)

# Change the date column to Type Date for summary statistics
data_list = lapply(data_list, mutate, date = as.Date(date))

# Create a variable that holds the current date in "Month Day" format to add to document text
today = format(Sys.Date(), "%B %d") 

# Filter temperature data to only include days that mach the current day and month
clean_data = lapply(data_list, GetDayTemp)

stpaul_data = cbind(clean_data[[1]], "Location" = rep("St. Paul, MN", nrow(clean_data[[1]])))

urbana_data = cbind(clean_data[[2]], "Location" = rep("Urbana, IL", nrow(clean_data[[2]])))

full_data = full_join(stpaul_data, urbana_data)

```

### Non presentation ready table
```{r}
# Create a basic summary table
table = data.frame(format(summary(clean_data[[1]]$TMAX), digits = 1), 
               format(summary(clean_data[[2]]$TMAX), digits = 1))
colnames(table) = c("St. Paul", "Urbana")

table
```

### Presentation ready table
```{r}
# Create a table to show presentation ready example
kable(table,
      caption = paste0("Summary statistics for historical average daily maximum temperatures on ", today, " in degrees Fahrenheit"),
      align = "r")
```

### Non presentation ready linear model summary
```{r}
## Fit a linear model to look at outputting summaries
fit = lm(TMIN ~ TMAX, data= stpaul_data)

# Basic summary call
summary(fit)
```

### Presentation ready linear model summary
```{r}
# Linear model summary using combination of kable and tidy for well formated table
kable(tidy(fit))
```

### Non presentation ready histogram
```{r, echo = FALSE}
# Create a histograms for an input location with basic ggplot call
hist_MN = ggplot(clean_data[[1]]) +
    geom_histogram(aes(TMAX), bins = 15) 

hist_IL = ggplot(clean_data[[2]]) +
    geom_histogram(aes(TMAX), bins = 15) 


cowplot::plot_grid(hist_MN, hist_IL)
```

### Presentation ready histogram
```{r, echo = FALSE}
# Creating a custom theme to use with histograms
theme_custom = function() {
  t = theme_classic() +
    theme(plot.title = element_text(hjust = 0.5),
            legend.position = "bottom") 
}

# Create a histogram with facets for comparison of St. Paul and Urbana temps
ggplot(full_data, aes(TMAX, after_stat(density))) +
  facet_wrap(~Location) +
  geom_histogram(binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3)), 
                 alpha = .5, 
                 fill = "#DC9D7F") +
  theme_custom() +
  labs(title = paste0("Historical Maximum Temperatures on ", today), caption = "Data Source: NOAA") +
  ylab("Density") +
  xlab("Maximum Temperature") 
  
```





